---
title: "Generate prediction points"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generate prediction points}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "2024-01-15"
author: "Insang Song"
---

```{r init-setting, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r load-terra}
library(terra)
```

# Objective
This vignette will demonstrate how the prediction grid points at 1-km
resolution are generated from the polygon data of the mainland US with
`terra` package.

# Strategy
- We set the upper left and lower right corners then make regular grid
points at 1,000 meter interval.
- `EPSG:5070`, Conus Albers equal area projection, is used throughout
this vignette.

```{r load-usmain}
usmain <- terra::vect("../tests/testdata/US-mainland-boundary.gpkg")
usmain <- terra::project(usmain, "EPSG:5070")
plot(usmain)

```

# Generate
Regular or random points can be generated from an extent or a polygon
object with `terra::spatSample()` or `sf::st_sample()`.
A faster way of generating regular points is to leverage a raster object,
where cells are organized in a regular grid. The code block below
generates 1-km resolution grid points following steps:

1. Identify corners to generate a rectangular extent
(i.e., `SpatExtent` object from `terra::ext()`)
2. Create a `SpatRaster` object with a fixed resolution and
coordinate system (in this case, `EPSG:5070`)
3. Assign a value to the void raster
4. Crop the raster object with the mainland US polygon
5. Convert the cropped raster to points (we have a `SpatVector` object)
6. Convert the `SpatVector` object to a three-column `data.frame` object
7. Save the `data.frame` object from step 6 as an RDS file

Steps 6 and 7 reduce the file size substantially as all data in the
`data.frame` from step 6 are in numeric type.
This means the data can be compressed efficiently.

```{r gen-grid-1km}
corner_ul <- c(-2.40, 3.26) * 1e6
corner_lr <- c(2.40, 0.12) * 1e6

corners <- c(corner_ul, corner_lr)
# reorganize xmin, ymin, xmax, ymax, which are ll, ur form
corners_re <- corners[c(1, 3, 4, 2)]
names(corners_re) <- c("xmin", "xmax", "ymin", "ymax")
corners_ext <- terra::ext(corners_re)
corners_ras <-
  terra::rast(
    corners_ext,
    resolution = c(1000L, 1000L),
    crs = "EPSG:5070"
  )

terra::values(corners_ras) <- 1L
corners_ras_sub <-
  terra::crop(
    corners_ras,
    usmain,
    snap = "out",
    mask = TRUE
  )

corners_pnts <- terra::as.points(corners_ras_sub)
corners_pnts_df <- as.data.frame(corners_pnts, geom = "XY")
corners_pnts_df$site_id <- seq(1, nrow(corners_pnts_df))
names(corners_pnts_df)[2:3] <- c("lon", "lat")
corners_pnts_df <- corners_pnts_df[, c("site_id", "lon", "lat")]

```

```{r save-rds, eval = FALSE}
saveRDS(
  corners_pnts_df,
  file = "./input/prediction_grid.rds",
  compress = "xz"
)

```

Below is a map of 10-km grid points in the mainland US for faster rendering.
The actual 1-km result will look denser.
```{r gen-grid-10km, echo = FALSE, message = FALSE, error = FALSE, fig.width = 8, fig.height = 4.8}
corners_ras10 <-
  terra::rast(
    corners_ext,
    resolution = c(10000L, 10000L),
    crs = "EPSG:5070"
  )

terra::values(corners_ras10) <- 1L
corners_ras_sub10 <-
  terra::crop(
    corners_ras10,
    usmain,
    snap = "out",
    mask = TRUE
  )

corners_pnts10 <- terra::as.points(corners_ras_sub10)
```

```{r plot-grid-10km}
plot(
  corners_pnts10,
  cex = 0.1,
  main = "10-km grid points in the mainland US"
)
```