---
title: "Create testdata on a small area"
output: html_document
date: "2024-01-02"
---

# Testdata area and period

### Small area

```{r}
nc_file <- system.file("shape/nc.shp", package = "sf")
nc <- terra::vect(nc_file)
sp_samp <- nc[nc$NAME %in% c("Durham", "Wake", "Orange")]
terra::plot(sp_samp)
```

### 2022-01-01 to 2022-01-02

```{r}
start_date <- as.Date("2022-01-01")
end_date <- as.Date("2022-01-02")
```

### Load covariate file names

```{r}
dict_path <- "../inst/extdata/downloaded_files_metadata.csv"
dict <- data.table::fread(dict_path, sep = ",")
```

### File path name

```{r}
string <- "/folder/yyyy/yday/varname_yyyymmdd.nc"
date <- as.Date("2022-08-01")
replace_dateinfo(string, date)
```

# Extract sampled area from each downloaded data

## NLCD

Extraction

```{r, eval = FALSE}
testdata_nlcd()
```

Example of output

```{r}
nlcd_sub_2019 <- terra::rast(
  "../tests/testdata/raw/nlcd/nlcd_2019_land_cover_l48_20210604.tif")
nlcd_sub_2021 <- terra::rast(
  "../tests/testdata/raw/nlcd/nlcd_2021_land_cover_l48_20230630.tif")
terra::plot(nlcd_sub_2019)
terra::plot(nlcd_sub_2021)
```

## Ecoregions

Extraction

```{r, eval = FALSE}
testdata_ecoregions()
```

Example of output

```{r}
ecor_sub <- terra::vect(
  "../tests/testdata/raw/ecoregions/us_eco_l3_state_boundaries.shp")
terra::plot(ecor_sub)
```

## NARR reanalysis

#### Variables stored as monthly rasters

Extraction

```{r, eval = FALSE}
narr_names <- c(
  "air_sfc", "albedo", "apcp", "dswrf", "evap", "hcdc", "hpbl",
  "lcdc", "lhtfl", "mcdc", "pr_wtr", "prate", "pres_sfc", "shtfl",
  "snowc", "soilm", "tcdc", "ulwrf_sfc", "uwnd_10m", "vis",
  "vwnd_10m", "weasd"
)
for (n in narr_names) {
  testdata_narr(var = n)
}
```

Example of output

```{r}
albedo_sub <- terra::rast("../tests/testdata/raw/narr/albedo.2022.nc")
terra::plot(albedo_sub)
```

#### Variables stored as monthly * atm levels rasters

Extraction

```{r, eval = FALSE}
narr_lev_names <- c("omega", "shum")
for (n in narr_lev_names) {
  testdata_narr_lev(var = n)
}
```

Example of output

```{r}
omega_sub <- terra::rast("../tests/testdata/raw/narr/omega.202201.nc")
terra::plot(omega_sub)
```

## GEOS CF

### AQC

Extraction

```{r, eval = FALSE}
aqc_path <- "../tests/testdata/raw/geos/aqc_tavg_1hr_g1440x721_v1/"
testdata_geos(testdata_path = aqc_path, collection = "aqc_tavg")
```

New file is not exactly the same than raw file but contains all covariates and units:

```{r}
r_path <- paste0(
  aqc_path,
  "GEOS-CF.v01.rpl.aqc_tavg_1hr_g1440x721_v1.20220101_0030z.nc4"
)
aqc_sub <- terra::rast(r_path)
aqc_sub
terra::plot(aqc_sub)
```

### CHM

Extraction

```{r}
chm_path <- "../tests/testdata/raw/geos/chm_tavg_1hr_g1440x721_v1/"
testdata_geos(testdata_path = chm_path, collection = "chm_tavg")
```

Example extracted

```{r}
r_path <- paste0(
  chm_path,
  "GEOS-CF.v01.rpl.chm_tavg_1hr_g1440x721_v1.20220101_0030z.nc4"
)
chm_sub <- terra::rast(r_path)
chm_sub
terra::plot(chm_sub[[1:16]])
terra::plot(chm_sub[[17:32]])
terra::plot(chm_sub[[33:48]])
terra::plot(chm_sub[[48:52]])
```

## TRI

TODO 

## AQS observations

```{r}
aqs_path <- paste0(
  dict[name == "aqs", folder],
  dict[name == "aqs", filename]
) |>
  replace_dateinfo(date = start_date)

aqs <- data.table::fread(aqs_path) |>
  subset(`Date Local` >= start_date & `Date Local` <= end_date)

aqs_vect <- terra::vect(aqs,
  geom = c("Longitude", "Latitude"),
  crs = "EPSG:4326"
)
aqs_samp <- sp_subset(aqs_vect) |>
  data.table::as.data.table()

data.table::fwrite(
  aqs_samp,
  "../tests/testdata/raw/durham_wake_orange_aqs.csv"
)
```
